\textbf{Political Instability and Financial Markets}

### 2018-07-12 doc/paper.tex
Added additional literature to literature review
Re-runs analysis with new data

\singlespacing
\begin{abstract}
We examine the relationship between political instability and daily returns of national stock indices. A common hypothesis is that political instability discourages investment and reduces economic growth. By contrast, we find that while financial volatility increases dramatically following ``irregular'' regime changes, different types of regime change have disparate effects on stock returns. We use daily financial data and a constant mean return event study model to show that abnormal returns following resignations are large and positive (6\%), while abnormal returns following assassinations are negative and smaller in magnitude (1\%). The impact of coups tends to be negative, but some events result in positive abnormal returns of 20\% or more. We also find that volatility increases during times of protest preceding resignations, but that no clear directionality is present. We therefore find that the type of political event and its expected impact on economic policy determines the direction of abnormal returns.
\end{abstract}
\doublespacing

@@

A common hypothesis is that political instability discourages investment and reduces economic growth by increasing policy uncertainty. This hypothesis is supported by cross country empirical evidence that political instability - as measured by regime change frequency or political violence - is negatively correlated with investment, GDP growth, and financial development in cross country regressions \citep{alesina1996political,alesina1996income,roe2011political}.\footnote{In particular, \citet{alesina1996political} show that countries have lower growth in years in which there are major government changes such as coups. On the other hand, \citet{londregan1990poverty} finds that coups have no effect on economic growth.} However, such hypotheses are overly simplistic. Not all political instability is equivalent - resignations, assassinations, and coups are each examples of political instability, but may have disparate effects on financial returns. Further, political instability may be seen as a \textit{positive} event if the current regime is regarded as strongly anti-business or anti-global.  \citet{alesina1996political} takes the possibility of a new government adopting better economic policies seriously, but argues that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions.

@@

We find that financial volatility increases dramatically following (and just before) coup d'etat, assassinations, and resignations. However, unlike the previously mentioned studies, we find that these irregular regime changes have disparate effects on the direction of stock returns. Abnormal returns following resignations are large and positive (6\%) while abnormal returns following assassinations are negative and smaller in magnitude (1\%). The impact of coups tends to be negative, but some events result in positive abnormal returns of 20\% or more. This finding is consistent with research that shows that ``good coups'' - coups that lead to democratization and economic liberalization - are not the norm and therefore on average coups should not be expected to lead to increased economic development and positive market returns \citep{derpanopoulos2015coups, marinov2014coups, powell2011global, thyne2016coup, varol2011democratic}. However, while ``good coups'' may be the minority, they do exist, and coups that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo.

@@

The second looks at how the variance of stock market returns reacts to political events. \citet{leblang2005government} show that right-wing governments increase the mean and volatility of stock prices in the United States and Britain. \citet{jensen2005market} find that a rise in the popularity of a Brazilian presidential candidate with uncertain future policies increased the volatility (but not the mean return) of the Brazilian stock market.

@@

Financial data are from the Global Financial Data database, which includes the longest available time series of stock prices. We collect data on national equity indices and two global equity indices, the S\&P/IFC Emerging Market Investable Composite and the Morgan Stanley Capital International (MSCI) World Index. The S\&P/IFC index includes securities from emerging markets while the MSCI index includes securities from developed markets only. We collect stock index data on every country in which there was a coup or coup attempt, an assassination or failed assassination, or a forced resignation.\footnote{The list of failed assassinations are from \citet{jones2009hit}. Coup attempts are those in category 2 in the CSP Coup d'etat dataset.} The longest available time series for these stock indices are listed in \autoref{tab:stock list}.

\singlespacing
\footnotesize
\begin{center}
\begin{longtable}[!ht]{l@{\extracolsep{\fill}}lll}
\caption{List of stock indices}\label{tab:stock list}\\
\vspace{-5pt}\\
\hline
\hline
\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Stock Index}&\multicolumn{1}{c}{Begin Date}&\multicolumn{1}{c}{End Date}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\hline
\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Stock Index}&\multicolumn{1}{c}{Begin Date}&\multicolumn{1}{c}{End Date}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
Argentina & Beunos Aires SE General Index & Dec-66 & Jan-13\\
Australia & Australia ASX All Ordinaries & Jan-58 & Jan-13\\
Bangladesh & Dhaka SE Index & Jan-90 & Apr-09\\
Canada & Canada S\&P/TSX 300 Composite & Jan-76 & Jan-13\\
Chile & Santiago IGBC General Index & Jan-75 & Jan-13\\
Colombia & Colombia IGBC General Index & Jan-92 & Jan-13\\
Ecuador & Ecuador Bolsa de Valores de Guayaquil (BVG) & Jan-94 & Nov-11\\
Egypt & Cairo SE Index & Dec-92 & Jan-13\\
Emerging Market & S\&P/IFC Emerging Markets Investable Composite & Jul-95 & Jan-13\\
France & France CAC All-Tradable Index & Sep-68 & Jan-13\\
Greece & Athens SE General Index & Oct-88 & Jan-13\\
India & Bombday SE SENSEX & Apr-79 & Jan-13\\
Indonesia & Jakarta SE Composite Index & Apr-83 & Jan-13\\
Iran & Tehran SE Price Index & Jan-95 & Mar-12\\
Israel & Tel Aviv 100 Index & May-87 & Jan-13\\
Japan & Tokyo SE Price Index (TOPIX) & Jan-53 & Jan-13\\
Latin America & Dow Jones Latin America Index & Jan-92 & Jan-13\\
Lithuania & Lithuania Lit-10 Index & Jan-99 & May-05\\
Malaysia & Malaysia KLSE Composite & Jan-80 & Jan-13\\
Nepal & Nepal NEPSE Stock Index & Jul-97 & Jul-99\\
Netherlands & Netherlands All-Share Price Index & Jan-80 & Jan-13\\
Pakistan & Karachi SE 100 Index & Jan-89 & Jan-13\\
Paraguay & Asuncion SE PDV General Index & Oct-93 & Sep-08\\
Peru & Lima SE General Index & Jan-82 & Jan-13\\
Philippines & Manila SE Composite Index & Jan-86 & Jan-13\\
Portugal & Oporto PSI-20 Index & Jan-86 & Jan-13\\
Singapore & Singapore FTSE ST Index & Jul-65 & Jan-13\\
South Korea & Korea SE Stock Price Index  & Jan-62 & Jan-13\\
Southeast Asia & Dow Jones Southeast Asia Index & Jan-92 & Jan-13\\
Spain & Madrid SE General Index & Aug-71 & Jan-13\\
Sri Lanka & Colombo SE All-Share Index & Dec-84 & Jan-13\\
Sweden & Sweden OMX Affarsvarlden General Index & Jan-80 & Jan-13\\
Taiwan & Taiwan SE Capitalization Weighted Index & Jan-67 & Jan-13\\
Thailand & Thailand SET General Index & Apr-75 & Jan-13\\
Tunisia & Tunisia SE Index & Dec-97 & Jan-13\\
Turkey & Instanbul IMKB 100 Price Index & Oct-87 & Jan-13\\
Ukraine & Ukraine PFTS OTC Index & Jan-98 & Jan-13\\
United Kingdom & UK FTSE All-Share Index & Nov-62 & Jan-13\\
United States & Dow Jones Industrial Average & Feb-1885 & Jan-13\\
Uruguay & Bolsa de Valores de Montevideo Index & Jan-08 & Jan-13\\
Venezuela & Dow Jones Venezuela Stock Index & Jan-92 & Jul-07\\
Venezuela & Caracas SE General Index & Jan-94 & Jan-13\\
World & MSCI World Price Index & Jan-76 & Jan-13\\
Zambia & Lusaka SE Index & Jan-02 & Apr-06\\
Zambia & Zambia LuSE Index & Jan-02 & Apr-06\\
\hline
\hline
\end{longtable}
\end{center}
\doublespacing
\normalsize

@@

\subsubsection{Coups} \label{subsec: Coups}
\autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat. \autoref{tab:AR-coups} contains all coups presented in \autoref{tab:coup list} with the exception of the partial/failed Venezuelan coup attempt of 2002. We examine this event as well, but do not include it in the sample in \autoref{tab:AR-coups} as it does not match our definition of a successful coup.

@@

The failed coup attempt against Hugo Chavez is a perfect natural experiment because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.

\autoref{fig:AR-Ven} provides graphical evidence on the effect of the coup attempt. The top panel shows CARs for the 10 days prior to and following the event, along with 95\% confidence intervals. The daily ARs and corresponding confidence intervals are displayed in the bottom panel. The abnormal return on April 12, the first trading day in which investors could react to the coup, was 19.5\%. The market reacted similarly, albeit in the opposite direction, to Chavez's reinstatement as president: the abnormal return on the next trading day, April 15, 2002, was -14\%.

The results in the figure are particularly striking given the discrepancy between the ARs on event days 0 and 1 and all other days. The only other day in which the ARs are statistically different from zero is on the third trading day after the coup attempt.

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those table \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven almost entirely by three events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

@@

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination is typically a random event that may not necessarily cause immediate change in economic policy. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 6\% and event day abnormal returns are only negative and statistically significant at even the ten percent level in one out of the fifteen resignations (Tunisia on June 31, 2011). However, these positive returns are short lived: the 7 day cumulative abnormal return is nearly identical to the event day abnormal return and the 20 day cumulative abnormal return is not statistically different from zero.

These results are also consistent with the idea that the expected impact of a political event on economic policy is highly important to the directionality of returns. The positive event day abnormal return following resignations is not surprising since resignations occurred because of ``poor performance and/or loss of authority.'' Leaders were often ousted following corruption charges, financial crises, and/or political violence. For example, consider Suharto's resignation from office as the President of Indonesia on May 1998, just after the 1997 Asian Financial Crisis.

Prior to the crisis, the Indonesian economy suffered from a number of ailments including crony capitalism, poor corporate governance, growing short-term debt and a corrupt banking system \citep{radelet2000onset}. In addition, Suharto's ill health created considerable policy uncertainty, which only added to the financial panic in Indonesia. Indonesia was also hesitant to adopt structural reforms which further weakened its economic image. As a result, the crisis gave additional ammunition to opposition leaders, and public opposition culminated in riots in mid-May 1998. Suharto eventually lost his military support and was forced to resign.

@@

While the Egyptian Revolution may have had positive pro-democracy effects on the Egpytian political system, its short-term impact on the economy was disastrous. As shown in \autoref{fig:CAR-Egypt}, abnormal returns on the Egyptian Stock Exchange Index (EGX 30) were around -7\% on January 26th and -10\% the day after. To prevent further decline during the uprising, the Egyptian Stock Exchanged closed at the end of trading on January 27th. President Mubarak resigned on February 11, but the market remained closed until March 23, when CARs declined by another 9\%, before rebounding slightly thereafter.

@@

As shown in \autoref{tab:non-parametric}, the mean event day abnormal returns for coups, assassinations and resignations are all statistically different from zero at the 5\% level using the rank test statistic and the abnormal returns for coups and resignations are significant at at least the 10\% level using the sign test. In addition, abnormal absolute returns for all events are statistically significant at the 1\% level using both the rank and sign statistics. On the other hand, the event day abnormal returns for the control portfolio are never statistically different from zero at even the 10\% level. Finally, the difference in means between the regime change country and the control portfolio are statistically different from zero for coups, resignations and all events combined when using two-sided p-values from the Wilcoxon rank test.\footnote{The Wilcoxon rank test is a non-parametric statistical technique that can be used to compare differences between matched samples.} In sum, these results suggests that the results from section~\autoref{subsec: Abnormal Returns} are not an artifact of deviations from normality or confounding world events.

The third concern, which is that the political events in this paper are not unexpected, is addressed in a number of ways. First, one can note that neither the 7-day or the 20-day mean CAR from tables \autoref{tab:AR-coups}, \autoref{tab:AR-ass} or \autoref{tab:AR-resignations} are statistically different than zero at even the 10\% level. Although some of the irregular regime changes have statistically significant pre-event CAR's, most of them do not, and the ones that do not always move in the same direction. For example, there was a negative 11.8\% 7-day CAR preceding the resignation of President Estrada of the Philippines in January of 2001 and a positive 14.7\% 7-day CAR preceding the resignation of President Fernando de la Rua of Argentina in December of 2001, even though both resignations were associated with large positive event day abnormal returns.

@@

Finally, to check whether our results are sensitive to model specification, we estimated market models in addition to the baseline constant returns model. We used the MSCI World Index as a proxy for their market return. [To do: add table for this]

@@

The results are consistent with the idea that government competence is a random walk and that government changes have large impacts on investor confidence. Financial volatility surrounding regime changes, which is often characterized by large positive \textit{and} negative stock returns, suggests that irregular regime changes increase policy uncertainty. The variation in the direction of abnormal returns is consistent with the idea that not all types of regime changes are equivalent in terms of expected economic policy outcomes. Abnormal returns are likely positive following resignations because those leaders were likely to be ``bad'' leaders, and coups may result in positive returns in the event of a transition to more democratic or pro-business leadership.

Although this study is based on only 35 events, the size of the effects suggests that it should not be dismissed. Irregular regime changes are important economic events than can drastically alter the economic landscape. For instance, the 21.26\% increase in the Venezuelan Stock Index following the temporary ousting of Hugo Chavez in 2002 is its 4th largest increase since 1994. In comparison, the Hong Kong Stock Index (Hang Seng Index) decreased by 21.7\% following the 1989 massacre of Chinese protestors in Tiananmen square and the Dow Jones Industrial Average in the United States fell by 22.61\% on Black Monday (1997), the largest one-day drop in its history.

@@

\pdfbookmark[1]{References}{References}
\bibliography{bibliography}

>>>>

\singlespacing
\begin{abstract}
We examine the relationship between political instability and daily returns of national stock indices. A common hypothesis is that political instability discourages investment and reduces economic growth. By contrast, we find that while financial volatility increases dramatically following ``irregular'' regime changes, different types of regime change have disparate effects on stock returns. We use daily financial data and a constant mean return event study model to show that abnormal returns following resignations are large and positive (4\%), while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The impact of coups tends to be negative (2\%), but some events result in positive abnormal returns of 10\% or more. We also find that volatility increases during times of protest preceding resignations, but that no clear directionality is present. We therefore find that the expected direction and magnitude of abnormal returns is dependent on the type of political event and its expected impact on economic policy.
\end{abstract}
\doublespacing

@@

A common hypothesis is that political instability discourages investment and reduces economic growth by increasing policy uncertainty. This hypothesis is supported by cross country empirical evidence that political instability - as measured by regime change frequency or political violence - is negatively correlated with investment, GDP growth, and financial development in cross country regressions \citep{alesina1996political,alesina1996income,roe2011political}.\footnote{In particular, \citet{alesina1996political} show that countries have lower growth in years in which there are major government changes such as coups. On the other hand, \citet{londregan1990poverty} finds that coups have no effect on economic growth.}

Such hypotheses are overly simplistic. Not all political instability is equivalent - resignations, assassinations, and coups are each examples of political instability, but may have disparate effects on financial returns. Further, political instability may be seen as a \textit{positive} event if the current regime is regarded as strongly anti-business or anti-global.  \citet{alesina1996political} takes the possibility of a new government adopting better economic policies seriously, but argues that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions.

@@

We find that financial volatility increases dramatically following (and just before) coup d'etat, assassinations, and resignations. However, unlike the previously mentioned studies, we find that these irregular regime changes have disparate effects on the direction of stock returns. Abnormal returns following resignations are large and positive (4\%) while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The impact of coups tends to be negative (2\%), but some events result in positive abnormal returns of 10\% or more. This finding is consistent with research that shows that ``good coups'' - coups that lead to democratization and economic liberalization - are not the norm and therefore on average coups should not be expected to lead to increased economic development and positive market returns \citep{derpanopoulos2015coups, marinov2014coups, powell2011global, thyne2016coup, varol2011democratic}. However, while ``good coups'' may be the minority, they do exist, and coups that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo.

@@

The second looks at how the variance of stock market returns reacts to political events. \citet{leblang2005government} show that right-wing governments increase the mean and volatility of stock prices in the United States and Britain. \citet{jensen2005market} find that a rise in the popularity of a Brazilian presidential candidate with uncertain future policies increased the volatility (but not the mean return) of the Brazilian stock market.  \citet{mcgillivray2003redistributive} examines the impact of electoral systems and partisan shifts on stock prices, but is primarily focused on inter-industry variation.

@@

Financial data are from the Global Financial Data database, which includes the longest available time series of stock prices. We collect data on national equity indices and two global equity indices, the S\&P/IFC Emerging Market Investable Composite and the Morgan Stanley Capital International (MSCI) World Index. The S\&P/IFC index includes securities from emerging markets while the MSCI index includes securities from developed markets only. We collect stock index data on every country in which there was a coup or coup attempt, an assassination or failed assassination, or a forced resignation.\footnote{The list of failed assassinations are from \citet{jones2009hit}. Coup attempts are those in category 2 in the CSP Coup d'etat dataset.} The longest available daily time series for these stock indices are listed in \autoref{tab:stock list}.

\singlespacing
\footnotesize
\begin{center}
\begin{longtable}[!ht]{l@{\extracolsep{\fill}}lll}
\caption{List of stock indices}\label{tab:stock list}\\
\vspace{-5pt}\\
\hline
\hline
\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Stock Index}&\multicolumn{1}{c}{Begin Date}&\multicolumn{1}{c}{End Date}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\hline
\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Stock Index}&\multicolumn{1}{c}{Begin Date}&\multicolumn{1}{c}{End Date}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
Argentina & Beunos Aires SE General Index & Dec-66 & Jan-17\\
Australia & Australia ASX All Ordinaries & Jan-58 & Jan-17\\
Bangladesh & Dhaka SE Index & Jan-90 & Jan-17\\
Canada & Canada S\&P/TSX 300 Composite & Jan-76 & Jan-17\\
Chile & Santiago IGBC General Index & Jan-75 & Jan-17\\
Colombia & Colombia IGBC General Index & Jan-92 & Jan-17\\
Ecuador & Ecuador Bolsa de Valores de Guayaquil (BVG) & Jan-94 & Jan-17\\
Egypt & Cairo SE Index & Dec-92 & Jan-17\\
Emerging Market & S\&P/IFC Emerging Markets Investable Composite & Jul-95 & Jan-17\\
Greece & Athens SE General Index & Oct-88 & Jan-17\\
India & Bombday SE SENSEX & Apr-79 & Jan-17\\
Indonesia & Jakarta SE Composite Index & Apr-83 & Jan-17\\
Iran & Tehran SE Price Index & Jan-95 & Jan-17\\
Israel & Tel Aviv 100 Index & May-87 & Jan-17\\
Japan & Tokyo SE Price Index (TOPIX) & Jan-53 & Jan-17\\
Latin America & Dow Jones Latin America Index & Jan-92 & Jan-17\\
Lithuania & Lithuania Lit-10 Index & Jan-99 & May-05\\
Malaysia & Malaysia KLSE Composite & Jan-80 & Jan-17\\
Nepal & Nepal NEPSE Stock Index & Jan-01 & Jan-17\\
Netherlands & Netherlands All-Share Price Index & Jan-80 & Jan-17\\
Pakistan & Karachi SE 100 Index & Jan-89 & Jan-17\\
Paraguay & Asuncion SE PDV General Index & Oct-93 & Sep-08\\
Peru & Lima SE General Index & Jan-82 & Jan-17\\
Philippines & Manila SE Composite Index & Jan-86 & Jan-17\\
Portugal & Oporto PSI-20 Index & Jan-86 & Jan-17\\
Singapore & Singapore FTSE ST Index & Jul-65 & Jan-17\\
South Korea & Korea SE Stock Price Index  & Jan-62 & Jan-17\\
Southeast Asia & Dow Jones Southeast Asia Index & Jan-92 & Jan-17\\
Spain & Madrid SE General Index & Aug-71 & Jan-17\\
Sri Lanka & Colombo SE All-Share Index & Dec-84 & Jan-17\\
Sweden & Sweden OMX Affarsvarlden General Index & Jan-80 & Jan-17\\
Taiwan & Taiwan SE Capitalization Weighted Index & Jan-67 & Jan-17\\
Thailand & Thailand SET General Index & Apr-75 & Jan-17\\
Tunisia & Tunisia SE Index & Dec-97 & Jan-17\\
Turkey & Instanbul IMKB 100 Price Index & Oct-87 & Jan-17\\
Ukraine & Ukraine PFTS OTC Index & Jan-98 & Jan-17\\
United Kingdom & UK FTSE All-Share Index & Nov-62 & Jan-17\\
United States & Dow Jones Industrial Average & Feb-1885 & Jan-17\\
Uruguay & Bolsa de Valores de Montevideo Index & Jan-08 & Jul-16\\
Venezuela & Dow Jones Venezuela Stock Index & Jan-92 & Jul-07\\
Venezuela & Caracas SE General Index & Jan-94 & Jan-17\\
World & MSCI World Price Index & Jan-76 & Jan-17\\
Zambia & Lusaka SE Index & Jan-02 & Apr-06\\
Zambia & Lusaka SE Index & Jul-11 & Jan-17\\

\hline
\hline
\end{longtable}
\end{center}
\doublespacing
\normalsize

@@

\subsubsection{Coups} \label{subsec: Coups}
\autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat. \autoref{tab:AR-coups} contains all coups presented in \autoref{tab:coup list} with the exception of the Argentinian coup of March 24, 1976 and the partial/failed Venezuelan coup attempt of 2002. The March 24, 1976 Argentinian coup is excluded from our analysis because the stock market remained closed from March 24 to April 5, 1976, or a period of twelve days.\footnote{Treating this twelve day period as a single day CARs results in a positive abnormal return of 58\%, a fluctuation that seems qualitatively unreasonable.}  We later examine the Venezuelan coup attempt of 2002 in detail, but do not include it in the sample in \autoref{tab:AR-coups} as it does not match our definition of a successful coup.

@@

The failed coup attempt against Hugo Chavez allows us to examine the effects of both pro-business and anti-business regime change separately from simple uncertainty because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.

\autoref{fig:AR-Ven} provides graphical evidence on the effect of the coup attempt. The top panel shows CARs for the 10 days prior to and following the event, along with 95\% confidence intervals. The daily ARs and corresponding confidence intervals are displayed in the bottom panel. The abnormal return on April 12, the first trading day in which investors could react to the coup, was 10\%. The market reacted similarly, albeit in the opposite direction, to Chavez's reinstatement as president: the abnormal return on the next trading day, April 15, 2002, was -8\%.

The results in the figure are particularly striking given the discrepancy between the ARs on event days 0 and 1 and all other days. The only days on which the ARs are statistically different from zero is on event days 0 and 1 after the coup attempt.

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those table \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

@@

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination is typically an unexpected event that may not necessarily cause immediate change in economic policy. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\% and event day abnormal returns are only negative and statistically significant at even the ten percent level in two out of the fifteen resignations (Pakistan on April 19, 1993 and Tunisia on June 31, 2011). However, these positive returns are short lived: the 7 day cumulative abnormal return is nearly identical to the event day abnormal return and the 20 day cumulative abnormal return is not statistically different from zero [[\textit{This doesn't look right to me. The positive returns seem persistent and to grow larger over time}]].

These results are also consistent with the idea that the expected impact of a political event on economic policy is highly important to the directionality of returns. The positive event day abnormal return following resignations is not surprising since resignations typically occurr because of poor performance and/or loss of authority. Leaders were often ousted following corruption charges, financial crises, and/or political violence.

For example, consider Muhammad Suharto's resignation from office as the President of Indonesia on May 1998, just after the 1997 Asian Financial Crisis. Prior to the crisis, the Indonesian economy suffered from a number of ailments including crony capitalism, poor corporate governance, growing short-term debt and a corrupt banking system \citep{radelet2000onset}. In addition, Suharto's ill health created considerable policy uncertainty, which only added to the financial panic in Indonesia. Indonesia was also hesitant to adopt structural reforms which further weakened its economic image. As a result, the crisis gave additional ammunition to opposition leaders, and public opposition culminated in riots in mid-May 1998. Suharto eventually lost his military support and was forced to resign.

@@

Despite heralding the beginning of the ``Arab Spring,'' the short-term impact of the Egyptian Revolution on the economy was disastrous. As shown in \autoref{fig:CAR-Egypt}, abnormal returns on the Egyptian Stock Exchange Index (EGX 30) were around -7\% on January 26th and -10\% the day after. To prevent further decline during the uprising, the Egyptian Stock Exchanged closed at the end of trading on January 27th. President Mubarak resigned on February 11, but the market remained closed until March 23, when CARs declined by another 9\%, before rebounding slightly thereafter.

@@

As shown in \autoref{tab:non-parametric}, the mean event day abnormal returns for coups, assassinations and resignations are all statistically different from zero at the 5\% level using the rank test statistic and the abnormal returns for coups, assassinations, and resignations are significant at at least the 10\% level using the sign test. In addition, abnormal absolute returns for all events are statistically significant at the 1\% level using both the rank and sign statistics. On the other hand, the event day abnormal returns for the control portfolio are never statistically different from zero at even the 10\% level [[\textit{This is not true for all events using the sign test, which seems odd to me}]]. Finally, the difference in means between the regime change country and the control portfolio are statistically different from zero for coups (1\% level), assassinations (10\% level), resignations (10\% level), and all events combined (1\% level) when using two-sided p-values from the Wilcoxon rank test.\footnote{The Wilcoxon rank test is a non-parametric statistical technique that can be used to compare differences between matched samples.} In sum, these results suggests that the results from section~\autoref{subsec: Abnormal Returns} are not an artifact of deviations from normality or confounding world events.

The third concern, which is that the political events in this paper are not unexpected, is addressed in a number of ways. First, one can note that neither the 7-day or the 20-day mean CAR from tables \autoref{tab:AR-coups}, \autoref{tab:AR-ass} or \autoref{tab:AR-resignations} are statistically different than zero at even the 10\% level [[\textit{Not true for resignations 20 day}]]. Although some of the irregular regime changes have statistically significant pre-event CAR's, most of them do not, and the ones that do not always move in the same direction. For example, there was a positive 14.7\% 7-day CAR preceding the resignation of President Fernando de la Rua of Argentina in December of 2001, even though the resignation was associated with large positive event day abnormal returns.

@@

Finally, to check whether our results are sensitive to model specification, we estimated market models in addition to the baseline constant returns model. We used the MSCI World Index as a proxy for their market return. [To do: add table for this]  [[\textit{Is this necessary?}]]

@@

The results are consistent with the idea that perceptions of government competence and changes in government  have large impacts on investor confidence. Financial volatility surrounding regime changes, which is often characterized by large negative \textit{and} positive stock returns, suggests that irregular regime changes increase policy uncertainty and volatility, but that uncertainty does not tell the whole story. The direction of returns depends on the expected outcome of the irregular regime change. The variation in the direction of abnormal returns is therefore consistent with the idea that not all types of regime changes are equivalent in terms of expected economic policy outcomes. Abnormal returns are likely positive following resignations because those leaders were likely to be ``bad'' leaders, and coups may result in positive returns in the event of a transition to more democratic or pro-business leadership.

\newpage

@@

\newpage
\pdfbookmark[1]{References}{References}
\bibliography{bibliography}
